---
title: "Prepare spatial data "
format: html
bibliography: references.bib
csl: ecology.csl
---

## Aim
- Generate a shapefile with the distribution of sinfonevada plots
- Select all plots that intersects with distribution of pine species in SN 
- Filter out data with forest cover > 50%

### Generate shapefile of Sinfonevada 

- Download `dicc_parcelas` from `diccionarios_sf_v2` and save as `dicc_parcelas.csv`. There is an issue: it contains 598 plots, but there is an unmacth with plots used by @GomezAparicioetal2009ArePine. We use the file `TODAS_LAS_VARIABLES_ambientales.xls` from @GomezAparicioetal2009ArePine  
- Download `dicc_fcc` from `diccionarios_sf_v2` and save as `dicc_fcc.csv`

```{r}
#| messsage: false
# Load required libraries
library(sf)
library(tidyverse)
library(here)
```


```{r}
all_plots <- readxl::read_excel(here::here("data/raw/lorena/TODAS_LAS_VARIABLES_ambientales.xls")) |> 
    dplyr::select(plot_id = ID_PARCELA, plot_code = NOMBRE, log = UTM_X_GPS, lat = UTM_Y_GPS) 

plot_sf <- st_as_sf(all_plots, coords = c("log", "lat"), crs = 23030) 

```

### Potential plots 

#### Pine plots from @GomezAparicioetal2009ArePine

```{r}
pine_lorena <- readxl::read_excel(here::here("data/raw/lorena/ALL sólo repoblaciones.xls")) |> 
  dplyr::select(plot_id = ID) |> 
  left_join(all_plots)
```


#### Pine plots from @NavarroGonzalezetal2013WeightLanduse 

```{r}
pine_irene <- readxl::read_excel(here::here("data/raw/irene/matriz_usos.xls")) |> 
  dplyr::select(plot_id = ID) |> 
  left_join(all_plots)
```

#### Read pine distribution 

- Select plots that intersect with pine distribution using REDIAM *Pinus* distribution maps. 

```{r}
#| eval: false

p <- "/Users/ajpelu/Google Drive/My Drive/_proyectos/pinares_andalucia/"

hal <- st_read(paste0(p, "pinus_halepensis20_2023.shp"), quiet = TRUE)
pin <- st_read(paste0(p, "pinus_pinaster_subsp20_2023.shp"), quiet = TRUE)
nig <- st_read(paste0(p, "pinus_nigra_subsp20_2023.shp"), quiet = TRUE)
syl <- st_read(paste0(p, "pinus_sylvestris_subsp20_2023.shp"), quiet = TRUE)


pinus <- bind_rows(hal, pin, nig, syl) |> st_transform(23030)

# read EENNPP 
sn <- st_read(here::here("data/geoinfo/sn_wdpa.shp"),
              quiet = TRUE) |> st_transform(23030)

pinus_sn <- st_intersection(pinus, sn)


pinus_sn <- st_make_valid(pinus_sn)
st_write(pinus_sn, here::here("data/geoinfo/pinares_sn_rediam.shp"), append = FALSE)

```


```{r}
pinares_sn_rediam <- st_read(here::here("data/geoinfo/pinares_sn_rediam.shp"), quiet = TRUE)
plot_sf_pinares_rediam <- st_filter(plot_sf, pinares_sn_rediam) 



plot_sf_pinares_rediam_aux <- st_join(plot_sf, pinares_sn_rediam) |> 
  dplyr::select(plot_id, plot_code, geometry, origen, pct_arbo, pct_sp, sp = especie) 


pct_arbo_df <- plot_sf_pinares_rediam_aux |>
  st_drop_geometry() |>
  group_by(plot_id) |>
  summarise(pct_arbo = unique(pct_arbo))



p <- plot_sf_pinares_rediam_aux |> 
  st_drop_geometry() |> 
  group_by(plot_id,plot_code) |> 
  mutate(rep = row_number()) |> 
  pivot_wider(
    id_cols = plot_id,
    names_from = rep,
    values_from = c(pct_sp, sp),
    names_sep = "_"
  ) |> 
  inner_join(pct_arbo_df)


pure <- p |> 
  filter(!is.na(pct_sp_1)) |> 
  filter(is.na(pct_sp_2)) |> 
  mutate(type = "monosp") |> 
  mutate(sp_princ = sp_1, 
         pct_sp_1 = as.character(pct_sp_1)) |> 
  rename(pct_sp = pct_sp_1, sps = sp_1) |> 
  dplyr::select(-c(sp_2, sp_3, pct_sp_2, pct_sp_3))


mix2 <- p |> 
  filter(!is.na(pct_sp_1)) |> 
  filter(!is.na(pct_sp_2)) |> 
  filter(is.na(pct_sp_3)) |> 
  mutate(sp_princ = ifelse(pct_sp_1 > pct_sp_2, sp_1, sp_2)) |> 
  unite(col = "sps", sp_1:sp_2, sep = " | ", remove = FALSE) |> 
  unite(col = "pct_sp", pct_sp_1:pct_sp_2, sep = " | ", remove = FALSE) |> 
  dplyr::select(-c(sp_1, sp_2, sp_3, pct_sp_2, pct_sp_1, pct_sp_2, pct_sp_3)) |> 
  mutate(type = "multi_sp")
  

mix3 <- p |> 
  filter(!is.na(pct_sp_1)) |> 
  filter(!is.na(pct_sp_2)) |> 
  filter(!is.na(pct_sp_3)) |> 
  rowwise() |> 
  mutate(pct = pmax(pct_sp_1, pct_sp_2, pct_sp_3),
           cuala = paste0("pct_sp_", which.max(c_across(c(pct_sp_1, pct_sp_2, pct_sp_3))))
  ) |> 
  ungroup() |> 
  mutate(cuala_aux = str_remove(cuala, "pct_"),
    sp_princ = case_when(
      cuala_aux == "sp_1" ~ sp_1,
      cuala_aux == "sp_2" ~ sp_2,
      cuala_aux == "sp_3" ~ sp_3
  )) |> 
  unite(col = "sps", sp_1:sp_3, sep = " | ", remove = FALSE) |> 
  unite(col = "pct_sp", pct_sp_1:pct_sp_3, sep = " | ", remove = FALSE) |> 
  dplyr::select(-c(sp_1, sp_2, sp_3, pct_sp_2, pct_sp_1, pct_sp_2, pct_sp_3, pct, cuala, cuala_aux)) |> 
  mutate(type = "multi_sp")

           
           

all_rediam <- bind_rows(pure, mix2, mix3)

plot_sf_pinares_rediam_unique <- plot_sf_pinares_rediam |> 
  left_join(all_rediam)


# Exportar esto 
# plot_sf_pinares_rediam_unique 
```


- Select plots that intersect with pine ecosystem using Ecosystem map of Sierra Nevada [@PerezLuqueetal2019MapEcosystems]


```{r}
sn_eco <- st_read(here::here("data/geoinfo/ecosistemas.shp"), 
                  quiet = TRUE) |> st_transform(23030)
# pinares_sn <- sn_eco |> 
#   filter(ECOSISTE_1 %in% c("Repoblaciones de coníferas", "Pinares autóctonos de P. sylvestris", "Pinares autóctonos sobre dolomías"))

plot_sf_pinares_ecosys <- st_join(plot_sf, sn_eco) |> 
  dplyr::select(plot_id, plot_code, ecosys_sn = ECOSISTE_1) 

```



#### Add FCC from Sinfonevada 

```{r}
# run only once 
dic_parcelas <- read_csv(here::here("data/raw/dicc_parcelas.csv"))

dic_fcc <- read_csv(here::here("data/raw/dicc_fcc.csv"))

plot_fcc <- dic_parcelas |> 
  dplyr::select(plot_id = id_parcela, fcc_arborea, cobertura_total) 
```


#### Elevation 
```{r}
plot_elev <- readxl::read_excel(here::here("data/raw/localizacion_parcelas.xls")) |> 
  dplyr::select(plot_id = ID_PARCELA, elev = ALTITUD) 
```



# All plots 
```{r}
data_all <- all_plots |> 
  left_join(
    pine_lorena |> mutate(lorena = 1)
  ) |> 
  left_join(
    pine_irene |> mutate(irene = 1)
  ) |> 
  left_join(
    plot_sf_pinares_rediam_unique |> mutate(pino_rediam = 1)) |> 
  left_join(plot_sf_pinares_ecosys |> 
              st_drop_geometry() |> 
              dplyr::select(plot_id, ecosys_sn)) |> 
  left_join(plot_fcc) |> 
  left_join(plot_elev) |> 
  st_drop_geometry()


data_all_sf <- st_as_sf(data_all, coords = c("log", "lat"), crs = 23030) 

# Export as shapefile 
sf::st_write(data_all_sf, here::here("data/geoinfo/parcelas_sinfonevada.shp"), append = FALSE) 

```



## Selection

```{r}
data_all_sf_lorena <- data_all_sf |> filter(lorena == 1)
sf::st_write(data_all_sf_lorena, here::here("data/geoinfo/parcelas_sinfonevada_lorena.shp"), append = FALSE) 
```



- In this [link](https://drive.google.com/drive/folders/0ByE6jT8_uLBRZVJtdWpmMDUyeGM?resourcekey=0-TMC3Cafd8vWEet7QTZ9gYg), there are several data about the paper of @GomezAparicioetal2009ArePine





